---
title: "Anti_vac Exploratory Data Analysis"
output: pdf_document
date: "2022-09-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(DataExplorer)
library(psych)
library(corrplot)
```

JL: From the original file antivac.csv, I created a new file antivac_renamed.csv, in which I manually shortened the variable names. For original variable names/survey items pls refer to the original dataset

```{r}
antivax <- read.csv(here("data", "antivac_renamed.csv"))
```

## Data cleaning 

```{r}
antivax <- antivax %>%
  select(- Collector.ID, -End_Date, -Consent)
```

```{r}
# renaming categorical labels 

antivax <- antivax %>%
  mutate(Highest_Education = case_when(
    Highest_Education == "Associates Degree (2 year college/university degree)" ~ "Associate_Degree",
    Highest_Education == "Bachelor Degree (4 year college/university degree)" ~ "Bachelor_Degree",
    Highest_Education == "Bachelor Degree (4 year college/university degree)" ~ "Bachelor_Degree",
    Highest_Education %in% c("Elementary school (grade level 1-8)", "No formal schooling") ~ "Elementary_or_less",
    Highest_Education == "High school (grade level 9-12/13)" ~ "High_school",
    Highest_Education == "Master's Degree" ~ "Master_Degree",
    Highest_Education == "Professional/Doctoral Degree (PhD, MD, DC etc..)" ~ "Professional_Degree"
  ))

antivax <- antivax %>%
  mutate(SM_influence = case_when(
    SM_influence == "I have not seen anything about vaccines on social media" ~ "Not_seen",
    SM_influence == "Influenced my opinion on vaccines (vaccines are better then I thought previously)" ~ "Influenced_better",
    SM_influence == "Influenced my opinion on vaccines (vaccines are worse then I thought previously)" ~ "Influenced_worse",
    SM_influence == "NOT influenced my opinion on vaccines" ~ "No_influence"
  ))

rename <- function(variable){
  variable <- ifelse(variable == "I believe this is NOT true", "False",
                     ifelse(variable == "I believe this is true", "True", "Unsure"))
  return(variable)
}

antivax[which(colnames(antivax) == "MMR_autism"):which(colnames(antivax) == "Protect_immunocompromised")]<- lapply(antivax[which(colnames(antivax) == "MMR_autism"):which(colnames(antivax) == "Protect_immunocompromised")], rename)
```

```{r}
# write.csv(antivax, here("data", "antivax_clean.csv"))
```

## Visualization 

```{r}
antivax_clean <- read.csv(here("data", "antivax_clean.csv"))
```

```{r message=F}
plot_bar(antivax_clean,
         ncol = 2)
```

## Converting outcome variable 

```{r}
correct_response <- c(F, F, F, F, F, F, T, T, T, T, T, T)
```

```{r}
score_convert <-function(variable, key){
  variable <- ifelse(variable == "True", 2,
                     ifelse(variable == "False", 0, 1))
  if (!key){
    variable <- 2-variable # reverse for items that have F as correct answer
  }
  
  return(variable)
}

```

```{r}

idx = 1

for (x in c(which(colnames(antivax_clean) == "MMR_autism"):which(colnames(antivax_clean) == "Protect_immunocompromised"))){
  antivax_clean[x] <- lapply(antivax_clean[x], score_convert, correct_response[idx])
  idx = idx + 1
}
```

## Looking at outcome variable 

```{r }
cormatrix <- cor(antivax_clean[which(colnames(antivax_clean) == "MMR_autism"):which(colnames(antivax_clean) == "Protect_immunocompromised")])

corrplot(cormatrix, method="color", tl.cex = .6, number.font = .4)
```


## PCA for key-dependent coded 

```{r}
variable <- antivax_clean[which(colnames(antivax_clean) == "MMR_autism"):which(colnames(antivax_clean) == "Protect_immunocompromised")]
```


```{r}
# PCA 
results <- prcomp(variable,scale = TRUE)

# R automatically returns the negative value. I apply -1 to get the original rotation value
results$rotation <- -1*results$rotation
results$rotation
```

```{r}
# R automatically returns negative value of how much each observation can be described by each PCA. Thus, I multiple th result by -1 to get the original value.
results$x <- -1*results$x
```

```{r}
# check how much variation each PC captures
# seems that PC1 and PC2 covers most 
results$sdev^2 / sum(results$sdev^2)
```

```{r}

#calculate total variance explained by each principal component
var_explained = results$sdev^2 / sum(results$sdev^2)
#create scree plot
qplot(c(1:12), var_explained) + 
  geom_line() + 
  xlab("Principal Component") + 
  ylab("Variance Explained") +
  ggtitle("Scree Plot") +
  ylim(0, 1)
```

```{r}
# visualize the cluster of PCs
plot_df = variable
colnames(plot_df) = c(1:12)
plot_re <- prcomp(plot_df, scale = TRUE)
plot_re$rotation <- -1*plot_re$rotation
plot_re$x <- -1*plot_re$x
plot(plot_re, data = plot_df,
         loadings = TRUE, loadings.colour = 'blue', loadings.label = TRUE, loadings.label.size = 4)
```

```{r}
fact = factanal(variable,factors=2)
fact
```

```{r}
library(factoextra)
fviz_pca_var(plot_re,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
```


## Chi squared test of independence

```{r}
table_hrs_vs_platform <- table(antivax_clean$Hours_use, antivax_clean$Social_media_used_most)
table_hrs_vs_platform_with_margins <- addmargins(table_hrs_vs_platform)
table_hrs_vs_platform_with_margins
chisq.test(table_hrs_vs_platform, simulate.p.value = TRUE)
```

## Modeling 

```{r}
alpha(variable)
```



```{r}
antivax_clean$score_sum <- rowSums(antivax_clean[c(which(colnames(antivax_clean) == "MMR_autism"):which(colnames(antivax_clean) == "Avoid_death"),25)]) # excluding data publicity, which PCA suggests dropping 
```

```{r}
ggplot(antivax_clean) +
  geom_histogram(aes(x = score_sum))
```

```{r}
vax_mod <- lm(score_sum ~ Hours_use + SM_influence + Highest_Education + SES + Social_media_used_most + Age + Gender + Continent + Trust_most, data = antivax_clean)
```

```{r}
summary(vax_mod)
```

```{r}
antivax_clean <- antivax_clean %>%
  mutate(hours_num = case_when(
    Hours_use == "0-2" ~ 1,
    Hours_use == "3-4" ~ 3.5,
    Hours_use == "5-6" ~ 5.5,
    Hours_use == "7-8" ~ 7.5,
    Hours_use == "9+" ~ 10
  ))
# make trust_most numeric
antivax_clean <- antivax_clean %>%
  mutate(trust_doc_gov = 1 * (Trust_most == "Doctors") + 1 * (Trust_most == "The government"),
         trust_other = 1 - trust_doc_gov
)
```

```{r}
vax_mod <- lm(score_sum ~ hours_num*SM_influence + Highest_Education + SES + Social_media_used_most + Age + Gender + trust_other, data = antivax_clean)
```

```{r}
summary(vax_mod)
```

# Lasso

```{r}
# make categorical variables numeric
antivax_clean_numeric <- antivax_clean %>% 
  mutate(SM_influence_worse = 1 * (SM_influence == "Influenced_worse"), # baseline = Influenced_better
         SM_influence_no_influence = 1 * (SM_influence == "No_influence"),
         SM_influence_not_seen = 1 * (SM_influence == "Not_seen"),
         Highest_Education_elementary_or_less = 1 * (Highest_Education == "Elementary_or_less"), # baseline = Associate_Degree
         Highest_Education_high = 1 * (Highest_Education == "High_school"),
         Highest_Education_bachelor = 1 * (Highest_Education == "Bachelor_Degree"),
         Highest_Education_master = 1 * (Highest_Education == "Master_Degree"),
         Highest_Education_professional = 1 * (Highest_Education == "Professional_Degree"),
         SES_middle = 1 * (SES == "Middle Class"), # baseline = Lower Class
         SES_upper = 1 * (SES == "Upper Class"),
         Social_media_used_most_Instagram = 1 * (Social_media_used_most == "Instagram"), # baseline = Facebook
         Social_media_used_most_Twitter = 1 * (Social_media_used_most == "Twitter"),
         Social_media_used_most_other = 1 * (Social_media_used_most == "Other (please specify)"),
         age = case_when(
          Age == "18-24" ~ 20,
          Age == "25-34" ~ 30,
          Age == "35-44" ~ 40,
          Age == "45-54" ~ 50,
          Age == "55-64" ~ 60,
          Age == "65+" ~ 70
        ),
        Gender_male = 1 * (Gender == "Male"), # baseline = Female
        Gender_other = 1 * (Gender == "Other / Prefer not to answer")
)
# create interaction variables
antivax_clean_numeric <- antivax_clean_numeric %>%
  mutate(hours_sum_SM_worse = hours_num * SM_influence_worse,
         hours_sum_SM_no_influence = hours_num * SM_influence_no_influence,
         hours_sum_SM_not_seen = hours_num * SM_influence_not_seen
)

# Lasso
library(glmnet)
attach(antivax_clean_numeric)

dependent_variable_vector <- score_sum
independent_variables_matrix <- cbind(1, hours_num,
                                      SM_influence_worse, SM_influence_no_influence, SM_influence_not_seen,
                                      Highest_Education_elementary_or_less, Highest_Education_high,
                                      Highest_Education_bachelor, Highest_Education_master, Highest_Education_professional,
                                      SES_middle, SES_upper, 
                                      Social_media_used_most_Instagram, Social_media_used_most_Twitter, Social_media_used_most_other,
                                      age, Gender_male, Gender_other,
                                      hours_sum_SM_worse, hours_sum_SM_no_influence, hours_sum_SM_not_seen,
                                      trust_other
                                )

cvfit <- cv.glmnet(independent_variables_matrix, dependent_variable_vector)
lambda.optimal <- cvfit$lambda.min
glmnet(independent_variables_matrix, dependent_variable_vector, lambda = lambda.optimal, alpha = 1)$beta

```













